Formalizing Social Engineering attacks in the Symbolic Model

### 2023-07-25 Seminar/article.tex
Finished the slides and fixed some spelling mistakes

\begin{itemize}
    \item \textit{Fresh} (groud) terms are prefixed by a \textasciitilde{} and model information that we can assume a party can check for freshness.
    \item \textit{Public} names are juxtaposed by a \$ and formalize ground information that any party within the network knows.
    \item \textit{Normal} terms are any form of term: fresh, public or built upon function symbols.
\end{itemize}

@@

We have seen how cryptographic operations can be formalized in Tamarin's syntax. Communication protocols involve far more complex dynamics than can not be expressed by atomic term manipulations, thus we introduce the idea of \textit{facts}.

In Tamarin, the execution of a protocol is modeled by the evolution of a multiset\footnote{Within this paper we define multisets through the $\{\{ ... \}\}$ notation: an empty multiset is represented as $\{\{\}\}$, while a multiset containing elements $x,y,z$ is represented as $\{\{x,y,z\}\}$} of facts (which represent the state of the system at any given moment). In knowledge representation, facts are "true" predicates that have a fixed arity and are composed of terms belonging to the supporting term-algebra; Tamarin follows this simple definition and requires the user to define them with a starting capital letter. In particular, there are two different versions of this construct that can be exploited during protocol modeling:

@@

\begin{definition}[Protocol rules]
    A A protocol rule is a multiset rewriting rule $l \xrightarrow{a} r$ such that
    \begin{enumerate}
        \item $l,a,r$ do not contain fresh names
        \item $l$ does not contain $\mono{K}$ and $\mono{Out}$ facts
        \item $r$ does not contain $\mono{Fr}$ and $\mono{K}$ facts
        \item The argument of any $\mono{Fr}$ fact belongs to set of fresh terms
        \item $r$ does not contain the function symbol $*$
        \item $l \xrightarrow{a} r$ satisfies:
        \begin{itemize}
            \item $vars(r) \subseteq vars(l) \cup \mathcal{V}_\textrm{pub}$
            \item $l$ only contains irreducible function symbols from the given signature or it is an instance if a rule that satisfies both conditions
        \end{itemize}
    \end{enumerate}
\end{definition}

@@

\begin{itemize}
    \item $[ \ ] \xrightarrow{\mono{Fresh}(~msg)} [ \mono{Fr}(~msg) ]$ allows for the generation of new fresh values.
    \item $[ \mono{Fr}(~msg) ] \to [ \mono{K}(~msg) ]$ allows for the generation of new fresh values by the attacker (expressed by the $\mono{K}$ fact)
    \item $[ \mono{Out}_{\textrm{ins}}(msg) ] \to [ \mono{K}(msg) ]$ allows the attacker to eavesdrop all messages travelling through the network. 
    \item $[ \mono{K}(msg) ] \xrightarrow{\mono{K}(msg)} [ \mono{In}_{\textrm{ins}}(msg) ]$ allows user to retrieve messages from the attacker-controlled network.
    \item $[ \ ] \to [ \mono{K}(\$x) ]$ allows the attacker to discover all public names.
    \item $[ \mono{K}(x_1,...,x_n) ] \to [ \mono{K}(\mono{f}(x_1,...,x_n)) ]$ allows the attacker to apply $n$-ary functions to arguments he already knows.
\end{itemize}

@@

\subsection{Including human knowledge in protocol formalization}

@@

\begin{figure}
\begin{alignat*}{3}
    &0.\  S &&: \textrm{knows}(R)\\
    &0.\  R &&: \textrm{knows}(S, m_2)\\
    &1.\  S \ins R &&: \textrm{fresh}(m_1).m_1\\
    &2.\  R \sec S &&: m_2
\end{alignat*}
\label{eq:toy_protocol}
\caption{Toy protocol in Alice and Bob notation}
\end{figure}

\begin{figure}
\begin{equation*}
    \langle \mono{Start}(S,R), \mono{Fresh}(S, m_1), \mono{Send}(S, \textrm{ins}, R, m_1), \mono{Receive}(S, \textrm{sec}, R, m_2) \rangle
\end{equation*}
\caption{Role script of role $S$ for the Toy protocol}
\end{figure}

@@

\begin{gather*}
    [ \ ] \xrightarrow{\mono{Start}(S,R)} [\mono{AgSt}(S,0,R)]\\
    [\mono{AgSt}(S,0,R), \mono{Fr}(m_1)] \xrightarrow{\mono{Fresh}(S,m_1)} [\mono{AgSt}(S,1,\langle R,m_1 \rangle)]\\
    [\mono{AgSt}(S,1,\langle R,m_1 \rangle)] \xrightarrow{\mono{Send}(S,\textrm{ins},R,m_1)} [\mono{AgSt}(S,2,\langle R,m_1 \rangle), \mono{Out}_\textrm{ins}(\langle S, R, m_1 \rangle)]\\
    [\mono{AgSt}(S,2,\langle R,m_1 \rangle), \mono{In}_\textrm{sec}(\langle R, S, m_2 \rangle)] \xrightarrow{\mono{Receive}(S, \textrm{sec}, R, m_2)} [\mono{AgSt}(S,3,\langle R, m_1, m_2 \rangle)]
\end{gather*}

Assuming we have the corresponding rules also for $R$, we can see that a succesfull execution of the protocols that follows the specification of \ref{eq:toy_protocol} will induce a trace that mirrors the role script of an agent (given that we ignore the action facts belonging to other parties' actions). Furthermore, note that the function symbols belonging to $\textrm{RoleActions}$ are now used as action facts: this is natural since in Tamarin we can only specify formulas (within lemmas and restrictions) over facts and not directly over terms.

@@

\begin{gather}
    [\mono{Fr}(x)] \xrightarrow{\mono{Fresh}(H, \langle t,x \rangle)} [\mono{!HK}(H, \langle t,x \rangle)] \label{eq:fresh_term_knowledge}\\
    [\mono{!HK}(H, \langle t, x \rangle)] \xrightarrow{\mono{Send}(H, \textrm{ins}, P, \langle t,x \rangle)} [\mono{Out}_{\textrm{ins}}(\langle H, P, \langle t,x \rangle \rangle)] \label{eq:telling_the_attacker}\\
    [\mono{In}_{\textrm{ins}}(\langle P, H, \langle t,x \rangle \rangle)] \xrightarrow{\mono{Receive}(H, \textrm{ins}, P, \langle t,x \rangle)} [\mono{!HK}(H, \langle t,x \rangle)] \label{eq:retrieving_from_the_attacker}
\end{gather}

@@

\begin{itemize}
    \item Rule \ref{eq:fresh_term_knowledge} allows to any human party to generate a fresh term and to add it to his knowledge, along with the relevant tag $t$. This models, for example, the generation of a new password by a registering user.
    \item Rule \ref{eq:telling_the_attacker} includes into the formalization the leakage of any information (both private and non) that makes up the human's knowledge database onto the network. A real-world example of this would be an involuntary password reveal by a user.
    \item Rule \ref{eq:retrieving_from_the_attacker} allows a human to update his existing knowledge based on new information received from the network. For example, an unexperienced user may receive a phishing mail explaining that he must re-enter his credentials into a new website because of security reasons.
\end{itemize}

>>>>

\begin{itemize}
    \item \textit{Fresh} (groud) terms are prefixed by a \textrm{\textasciitilde}{} and model information that we can assume a party can check for freshness.
    \item \textit{Public} names are juxtaposed by a \$ and formalize ground information that any party within the network knows.
    \item \textit{Normal} terms are any form of term: fresh, public or built upon function symbols.
\end{itemize}

@@

We have seen how cryptographic operations can be formalized in Tamarin's syntax. Communication protocols involve far more complex dynamics than the ones that can be expressed by atomic term manipulations, thus we introduce the idea of \textit{facts}.

In Tamarin, the execution of a protocol is modeled by the evolution of a multiset\footnote{Within this paper we define multisets through the $\{\{ ... \}\}$ notation: an empty multiset is represented as $\{\{\}\}$, while a multiset containing elements $x,y,z$ is represented as $\{\{x,y,z\}\}$} of facts (which represent the state of the system at any given moment). In logic, facts are predicates that have a fixed arity and are composed of terms belonging to the supporting term-algebra; Tamarin follows this simple definition and requires the user to define them with a starting capital letter. In particular, there are two different versions of this construct that can be exploited during protocol modeling:

@@

\begin{definition}[Protocol rules]
    A A protocol rule is a multiset rewriting rule $l \xrightarrow{a} r$ such that
    \begin{enumerate}
        \item $l,a,r$ do not contain fresh names
        \item $l$ does not contain $\mono{K}$ and $\mono{Out}$ facts
        \item $r$ does not contain $\mono{Fr}$ and $\mono{K}$ facts
        \item The argument of any $\mono{Fr}$ fact belongs to the set of fresh terms
        \item $r$ does not contain the function symbol $*$
        \item $l \xrightarrow{a} r$ satisfies:
        \begin{itemize}
            \item $vars(r) \subseteq vars(l) \cup \mathcal{V}_\textrm{pub}$
            \item $l$ only contains irreducible function symbols from the given signature or it is an instance of a rule that satisfies both conditions
        \end{itemize}
    \end{enumerate}
\end{definition}

@@

\begin{itemize}
    \item $[ \ ] \xrightarrow{\mono{Fresh}(\textrm{\textasciitilde} msg)} [ \mono{Fr}(\textrm{\textasciitilde} msg) ]$ allows for the generation of new fresh values.
    \item $[ \mono{Fr}(\textrm{\textasciitilde} msg) ] \xrightarrow{\mono{Fresh}(\textrm{\textasciitilde} msg)} [ \mono{K}(\textrm{\textasciitilde} msg) ]$ allows for the generation of new fresh values by the attacker (expressed by the $\mono{K}$ fact)
    \item $[ \mono{Out}_{\textrm{ins}}(msg) ] \to [ \mono{K}(msg) ]$ allows the attacker to eavesdrop all messages travelling through the network. 
    \item $[ \mono{K}(msg) ] \xrightarrow{\mono{K}(msg)} [ \mono{In}_{\textrm{ins}}(msg) ]$ allows user to retrieve messages from the attacker-controlled network.
    \item $[ \ ] \to [ \mono{K}(\$x) ]$ allows the attacker to discover all public names.
    \item $[ \mono{K}(x_1,...,x_n) ] \to [ \mono{K}(\mono{f}(x_1,...,x_n)) ]$ allows the attacker to apply $n$-ary functions to arguments he already knows.
\end{itemize}

@@

\subsection{Including human knowledge in protocol formalization}\label{sec:including_humans}

@@

\begin{figure}
\begin{alignat*}{3}
    &0.\  S &&: \textrm{knows}(R)\\
    &0.\  R &&: \textrm{knows}(S, m_2)\\
    &1.\  S \ins R &&: \textrm{fresh}(m_1).m_1\\
    &2.\  R \sec S &&: m_2
\end{alignat*}
\caption{Toy protocol in Alice and Bob notation}
\label{eq:toy_protocol}
\end{figure}

\begin{figure}
\begin{equation*}
    \langle \mono{Start}(S,R), \mono{Fresh}(S, m_1), \mono{Send}(S, \textrm{ins}, R, m_1), \mono{Receive}(S, \textrm{sec}, R, m_2) \rangle
\end{equation*}
\caption{Role script of role $S$ for the Toy protocol}
\label{eq:role_script}
\end{figure}

@@

\begin{figure}
\begin{gather*}
    [ \ ] \xrightarrow{\mono{Start}(S,R)} [\mono{AgSt}(S,0,R)]\\
    [\mono{AgSt}(S,0,R), \mono{Fr}(m_1)] \xrightarrow{\mono{Fresh}(S,m_1)} [\mono{AgSt}(S,1,\langle R,m_1 \rangle)]\\
    [\mono{AgSt}(S,1,\langle R,m_1 \rangle)] \xrightarrow{\mono{Send}(S,\textrm{ins},R,m_1)} [\mono{AgSt}(S,2,\langle R,m_1 \rangle), \mono{Out}_\textrm{ins}(\langle S, R, m_1 \rangle)]\\
    [\mono{AgSt}(S,2,\langle R,m_1 \rangle), \mono{In}_\textrm{sec}(\langle R, S, m_2 \rangle)] \xrightarrow{\mono{Receive}(S, \textrm{sec}, R, m_2)} [\mono{AgSt}(S,3,\langle R, m_1, m_2 \rangle)]
\end{gather*}
\caption{Rewriting rules for role script \ref{eq:role_script}}
\end{figure}
Assuming we have the corresponding rules also for $R$, we can see that a succesfull execution of the protocols that follows the specification of \ref{eq:toy_protocol} will induce a trace that mirrors the role script of an agent (given that we ignore the action facts belonging to other parties' actions). Furthermore, note that the function symbols belonging to $\textrm{RoleActions}$ are now used as action facts: this is natural since in Tamarin we can only specify formulas (within lemmas and restrictions) over facts and not directly over terms.

@@

\begin{figure}[b]
\begin{gather}
    [\mono{Fr}(\textrm{\textasciitilde}x)] \xrightarrow{\mono{Fresh}(H, \langle t,\textrm{\textasciitilde}x \rangle)} [\mono{!HK}(H, \langle t,\textrm{\textasciitilde} x \rangle)] \label{eq:fresh_term_knowledge}\\
    [\mono{!HK}(H, \langle t, x \rangle)] \xrightarrow{\mono{Send}(H, \textrm{ins}, P, \langle t,x \rangle)} [\mono{Out}_{\textrm{ins}}(\langle H, P, \langle t,x \rangle \rangle)] \label{eq:telling_the_attacker}\\
    [\mono{In}_{\textrm{ins}}(\langle P, H, \langle t,x \rangle \rangle)] \xrightarrow{\mono{Receive}(H, \textrm{ins}, P, \langle t,x \rangle)} [\mono{!HK}(H, \langle t,x \rangle)] \label{eq:retrieving_from_the_attacker}
\end{gather}
\caption{Rewriting rules for the untrained human}
\end{figure}

@@

\begin{itemize}
    \item Rule \ref{eq:fresh_term_knowledge} allows to any human party to generate a fresh term and to add it to his knowledge, along with the relevant tag $t$. This models, for example, the generation of a new password by a registering user. Note that the tag should be assigned through the $Tag$ function described in \ref{sec:including_humans} with $x$ as his argument, but here this detail was left out for clarity purposes.
    \item Rule \ref{eq:telling_the_attacker} includes into the formalization the leakage of any information (both private and non) that makes up the human's knowledge database onto the network. A real-world example of this would be an involuntary password reveal by a user.
    \item Rule \ref{eq:retrieving_from_the_attacker} allows a human to update his existing knowledge based on new information received from the network. For example, an unexperienced user may receive a phishing mail explaining that he must re-enter his credentials into a new website because of security reasons.
\end{itemize}
